For submission to the Journal of Integrated Pest Management

### 2021-03-26 ofv_presentation_2021.Rmd
New Rmd for creating ESA SEB 2021 presentation

---
title: "For submission to the Journal of Integrated Pest Management"
subtitle: "Category: 'Issues'"
output: 
  word_document:
    reference_docx: ../r-word-templates/word-template-jis.docx
# - html_document
bibliography: 'thesis.bib'
csl: 'j_med_entomol.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

@@

### Keywords:
False spider mite, flat mite, *Brevipalpus*-transmitted viruses, *Liriope*, Nolinoideae, *Ophiopogon*, Ruscaceae, Rutaceae, Asparagaceae, orchid, Orchidaceae, pests, ornamental plants, orchid fleck virus.

\pagebreak
*Orchid fleck virus* (OFV) is the type member for the genus *Dichorhavirus*, family Rhabdoviridae; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Other members of this genus are: *Citrus chlorotic spot virus*, *Citrus leprosis virus* N, *Clerodendrum chlorotic spot virus* and *Coffee ringspot virus* [@Dietzgen2018a]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the only known vector for dichorhaviruses [@Maeda1998], and *Brevipalpus californicus* (Banks) mites are the only mite known to transmit OFV in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms dependent on the infected plant species as well as the strain of the OFV [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`).
<!-- These lesions are thought to be associated with the feeding sites of the mite vector [@cite?]. Also, is it possibly a sign of RNAi plant defenses? See Hull Plant Virology -->

OFV was first described infecting *Cymbidium* orchids in Japan [@Doi1977]. Many countries have reported OFV and OFV-like rhabdoviruses infecting orchids worldwide [@Kondo2003], including Asia: [China [@Peng2017], Korea [@Chang1991, @Zheng2013]], Africa: [South Africa [@Blanchfield2001]], North America: [The United States [@Blanchfield2001; @Bratsch2015]], South America: [Brazil [@Kitajima1974; @Kitajima2001], Colombia [@Kubo2009b], Costa Rica [@FreitasAstua2002], Paraguay [@RamosGonzalez2015]], Europe: [Denmark [@Begtrup1972], France [@Sauvetre2018], Germany [@Lesemann1975; @Petzold1971]] and Oceania: [Australia [@Lesemann1971; @Lesemann1975; @Gibbs2000], Fiji [@Pearson1993], Vanuatu [@Pearson1993]]. The prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a].

OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], as well as Rutaceae: *Citrus* where it causes citrus leprosis-like symptoms [@Beltran-Beltran2020; @GarciaEscamilla2018; @Roy2020]. Mechanical transmission of OFV is possible under lab conditions to various Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, which includes a diverse array of various monocotyledonous lilliod plants which are native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* Ker Gawler (Asparagaceae: Nolinoidaea) are considered the most important ground cover sold by the nursery industry in southeastern US [@Mcharo2003].

Viral infections of suspected plant leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples tested negative for begomovirus, potyvirus, tospovirus as well as for Impatiens necrotic spot virus, Tobacco mosaic virus and Tomato spotted wilt virus. The infected materials were subsequently sent to the Florida Department of Agriculture and Consumer Services (FDACS). The presence of OFV was confirmed using OFV generic R2-Dicho-GF and R2-Dicho-GR primers [@Roy2020] by one step conventional RT-PCR, amplifying ~800 nt of L-gene (RNA2) amplicon from an infected *Liriope* leaf sample. The RT-PCR amplicon was sequenced using Sanger sequencing and confirmed the presence of OFV in the tested samples. Sanger sequencing of RT-PCR amplicons shared 98\% nucleotide identity with orchid strains of OFV (GenBank Accession numbers: AB244418 and LC222630) [@Kondo2006; @Kondo2017]. Further surveys of putatively OFV-infected plants were taken during subsequent visits to the initial site of collection. Plant samples included various cultivars of Nolinoid plants, including *Liriope* spp., *Ophiopogon* spp., and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) which were located at the same site as the original infected *Liriope* spp., cv 'Gigantea'. *A. elatior* infected with OFV have chlorotic leaves or chlorotic flecks (`r citef("fig_1")`). No ringspots have been observed on *A. elatior* to date. All plant samples were tested via RT-PCR at the NFREC<!-- (`r citet("tab_1")`) --> to confirm the presence of OFV.

Further surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in Leon county which were observed with phase contrast microscopy. We encountered both an unidentified eriophyoid mite species on the *Liriope* spp., along with flat mites on all species tested (*Liriope* spp., *Ophiopogon* spp., and *A. elatior*). The flat mites were originally identified as *Brevipalpus californicus* (Banks) sensu lato, an identity which was confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. *Brevipalpus* mites have been previously associated with OFV [@Beltran-Beltran2020; @GarciaEscamilla2018; @Dietzgen2018b] as well as similar diseases [@Kitajima2010] and are known to feed on a large variety of economically important plants [@Childers2003a; @Akyazi2017]. Unfortunately, the *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] whose identification can be improved with molecular methods as well as more advanced microscopy techniques, such as cryo scanning electron microscopy (currently known as Cryo-SEM) [@Skoracka2015; @Leon2010; @Beard2015]. With that in mind, we collected additional samples   of the mites from the same site as the original OFV detection and sent them to the USDA-ARS in Beltsville to observe the mites with Cryo-SEM  techniques, which agreed with both prior identifications of *B. californicus* s.l. <!-- (`r citef("fig_2")`) -->but revealed the presence of two other species: *B. obovatus* and *B. confusus*.

The first report of OFV in the United State is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids.  The significance of this report is their reference to spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, @Ko1985 made no mention of mites or further investigations of this virus. The first certain report of OFV in the US was made by @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids in the US, using TEM of ultrathin sections of plant tissue as well as molecular sequence analysis and its association with *Brevipalpus* mites. The authors did not make a conclusive species identification but suggested the mite species was *B. californicus* group as the vector as referred by @Kondo2003.

@@

### Acknowledgements
We would like to give a special thanks to the Tallahassee Museum for their patience, cooperation, and support with collecting plant samples. We also want to thank Drs. Sam Bolton, FDACS and Aline Tassi, Univ. of Sao Paulo, Brazil for checking the mites we have sent for species validation. We are especially indebted to the late Dr. Gary Bauchan for permitting us to use these beautiful LT-SEM images, he will be greatly missed. This research was partly funded by the USDA National Institute of Food and Agriculture, Hatch project FLA-NFC-005607. Mention of trade names or commercial products in this publication is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the USDA; USDA is an equal opportunity provider and employer.

### References
<div id="refs"></div>

>>>>

---
title: "For submission to the Journal of Integrated Pest Management"
subtitle: "Category: 'Issues'"
output: 
  beamer_presentation: 
    latex_engine: xelatex
bibliography: 'thesis.bib'
csl: 'j_med_entomol.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

@@

### Background
*Orchid fleck virus* (OFV) is the type member for the genus *Dichorhavirus*, family Rhabdoviridae; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Other members of this genus are: *Citrus chlorotic spot virus*, *Citrus leprosis virus* N, *Clerodendrum chlorotic spot virus* and *Coffee ringspot virus* [@Dietzgen2018a]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the only known vector for dichorhaviruses [@Maeda1998], and *Brevipalpus californicus* (Banks) mites are the only mite known to transmit OFV in a persistent propagative manner [@Kondo2003].

### Symptoms
OFV-infected plants exhibit various symptoms dependent on the infected plant species as well as the strain of the OFV [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`).
<!-- These lesions are thought to be associated with the feeding sites of the mite vector [@cite?]. Also, is it possibly a sign of RNAi plant defenses? See Hull Plant Virology -->

### Distribution
OFV was first described infecting *Cymbidium* orchids in Japan [@Doi1977]. Many countries have reported OFV and OFV-like rhabdoviruses infecting orchids worldwide [@Kondo2003], including Asia: [China [@Peng2017], Korea [@Chang1991, @Zheng2013]], Africa: [South Africa [@Blanchfield2001]], North America: [The United States [@Blanchfield2001; @Bratsch2015]], South America: [Brazil [@Kitajima1974; @Kitajima2001], Colombia [@Kubo2009b], Costa Rica [@FreitasAstua2002], Paraguay [@RamosGonzalez2015]], Europe: [Denmark [@Begtrup1972], France [@Sauvetre2018], Germany [@Lesemann1975; @Petzold1971]] and Oceania: [Australia [@Lesemann1971; @Lesemann1975; @Gibbs2000], Fiji [@Pearson1993], Vanuatu [@Pearson1993]]. The prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a].

### Infectivity
OFV naturally infects more than fifty species of Orchidaceae [@Kitajima2010; @Peng2013], some Asparagaceae (Nolinoidaea) [@Mei2016; @Dietzgen2018b], as well as Rutaceae: *Citrus* where it causes citrus leprosis-like symptoms [@Beltran-Beltran2020; @GarciaEscamilla2018; @Roy2020]. Mechanical transmission of OFV is possible under lab conditions to various Chenopodiaceae, Aizoaceae, Fabaceae, and Solanaceae [@Chang1976; @Kondo2003; @Peng2013].

### Detection
During June 2020, chlorotic ringspot symptoms were observed on Giant Lilyturf *Liriope* spp., cv. 'Gigantea' in a landscape of Leon County, Florida (`r citef("fig_1")`). *Liriope* belong to a group of plants in the family Asparagaceae, subfamily Nolinoidaea, which includes a diverse array of various monocotyledonous lilliod plants which are native to southeastern Asia [@Chase2009; @Meng2021]. *Liriope* and the closely related *Ophiopogon* Ker Gawler (Asparagaceae: Nolinoidaea) are considered the most important ground cover sold by the nursery industry in southeastern US [@Mcharo2003].

### Testing
Viral infections of suspected plant leaf samples were initially tested at the Plant Disease Diagnostic Clinic at the North Florida Research and Education Center (NFREC) in Quincy, FL. All the samples tested negative for begomovirus, potyvirus, tospovirus as well as for Impatiens necrotic spot virus, Tobacco mosaic virus and Tomato spotted wilt virus. The infected materials were subsequently sent to the Florida Department of Agriculture and Consumer Services (FDACS). The presence of OFV was confirmed using OFV generic R2-Dicho-GF and R2-Dicho-GR primers [@Roy2020] by one step conventional RT-PCR, amplifying ~800 nt of L-gene (RNA2) amplicon from an infected *Liriope* leaf sample. The RT-PCR amplicon was sequenced using Sanger sequencing and confirmed the presence of OFV in the tested samples. Sanger sequencing of RT-PCR amplicons shared 98\% nucleotide identity with orchid strains of OFV (GenBank Accession numbers: AB244418 and LC222630) [@Kondo2006; @Kondo2017]. Further surveys of putatively OFV-infected plants were taken during subsequent visits to the initial site of collection. Plant samples included various cultivars of Nolinoid plants, including *Liriope* spp., *Ophiopogon* spp., and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) which were located at the same site as the original infected *Liriope* spp., cv 'Gigantea'. *A. elatior* infected with OFV have chlorotic leaves or chlorotic flecks (`r citef("fig_1")`). No ringspots have been observed on *A. elatior* to date. All plant samples were tested via RT-PCR at the NFREC<!-- (`r citet("tab_1")`) --> to confirm the presence of OFV.

### Map
Further surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in both Leon and Alachua counties. In addition to the discovery of the OFV-infected plants, we collected mites from symptomatic plants in Leon county which were observed with phase contrast microscopy. We encountered both an unidentified eriophyoid mite species on the *Liriope* spp., along with flat mites on all species tested (*Liriope* spp., *Ophiopogon* spp., and *A. elatior*). The flat mites were originally identified as *Brevipalpus californicus* (Banks) sensu lato, an identity which was confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. *Brevipalpus* mites have been previously associated with OFV [@Beltran-Beltran2020; @GarciaEscamilla2018; @Dietzgen2018b] as well as similar diseases [@Kitajima2010] and are known to feed on a large variety of economically important plants [@Childers2003a; @Akyazi2017]. Unfortunately, the *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] whose identification can be improved with molecular methods as well as more advanced microscopy techniques, such as cryo scanning electron microscopy (currently known as Cryo-SEM) [@Skoracka2015; @Leon2010; @Beard2015]. With that in mind, we collected additional samples   of the mites from the same site as the original OFV detection and sent them to the USDA-ARS in Beltsville to observe the mites with Cryo-SEM  techniques, which agreed with both prior identifications of *B. californicus* s.l. <!-- (`r citef("fig_2")`) -->but revealed the presence of two other species: *B. obovatus* and *B. confusus*.

### Previous Reports
The first report of OFV in the United State is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids.  The significance of this report is their reference to spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, @Ko1985 made no mention of mites or further investigations of this virus. The first certain report of OFV in the US was made by @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids in the US, using TEM of ultrathin sections of plant tissue as well as molecular sequence analysis and its association with *Brevipalpus* mites. The authors did not make a conclusive species identification but suggested the mite species was *B. californicus* group as the vector as referred by @Kondo2003.

@@

### Acknowledgements
We would like to give a special thanks to the Tallahassee Museum for their patience, cooperation, and support with collecting plant samples. We also want to thank Drs. Sam Bolton, FDACS and Aline Tassi, Univ. of Sao Paulo, Brazil for checking the mites we have sent for species validation. We are especially indebted to the late Dr. Gary Bauchan for permitting us to use these beautiful Cryo-SEM images, he will be greatly missed. This research was partly funded by the USDA National Institute of Food and Agriculture, Hatch project FLA-NFC-005607. Mention of trade names or commercial products in this publication is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the USDA; USDA is an equal opportunity provider and employer.

<!-- ### References -->
<!-- <div id="refs"></div> -->
